### 2022-08-02 thesis/conclusion.tex
Major grammatical and formatting changes to third draft of thesis

\chapter{Conclusion}
\label{chap:conclusion}
In this thesis, we covered a specific technique to perform reachability analysis over non-linear dynamical systems by leveraging parallelotope bundles and a specific non-linear optimization problem whose solutions determine certain template directions' offsets required to express the reachable set.
%
We presented the original algorithm utilizing static template directions.
%
We then investigated two techniques for generating templates \emph{dynamically}: first using linear approximation of the dynamics, and second using PCA.
%
We demonstrated that these techniques improve the accuracy of reachable set by an order of magnitude when compared to static or random template directions.
%
We also observed that both these techniques improve the accuracy of the reachable sets for different benchmarks.
%
Finally, we experimented with potential applications of the Bernstein expansion technique to real-time reachability. We found that several hurdles and barriers arise to effectively utilizing the pruning method developed in the pre-computation phase.
%
Several remarks on ideas towards improvement of the methods we developed are listed below:
\begin{enumerate}
  \item Koopman linearization techniques for computing alternative linear approximation template directions with alternative optimization methods could yield interesting modifications to Algorithms \ref{alg:old} and \ref{alg:new} \cite{bak2021reachability}.

@@

  \item Step five of the pre-computation algorithm, where it simply takes the intervals which dominate all the other intervals, may be too crude to properly yield a proper speed-up. Are there any other methods of adding basis monomials to the lookup table which take into account the degree discrepancies outlined in Section \ref{sec:deg_disc}?
\end{enumerate}

>>>>

\chapter{Conclusion}
\label{chap:conclusion}
In this thesis, we covered a specific technique to perform reachability analysis over non-linear dynamical systems through leveraging parallelotope bundles. A particular non-linear optimization problem determines the offset of template directions required to express the reachable set.
%
We presented the original algorithm utilizing static template directions.
%
We then investigated two techniques for generating templates \emph{dynamically} and \emph{automatically}: the first using linear approximation of the dynamics, and the second using PCA.
%
We demonstrated that these techniques improve the accuracy of the reachable set of several benchmarks by an order of magnitude when compared to static or random template directions.
%
Finally, we experimented with potential applications of the Bernstein expansion technique to real-time reachability. We found that several hurdles and barriers arise to effectively utilizing the pruning method developed in the pre-processing phase.
%
Several remarks on ideas towards improvement of the methods we developed are listed below:
\begin{enumerate}
  \item Koopman linearization techniques for computing alternative linear approximation template directions with other optimization methods could yield interesting modifications to Algorithms \ref{alg:old} and \ref{alg:new} \cite{bak2021reachability}.

@@

  \item Simply taking the intervals which dominate all the other intervals may be too crude to properly yield an effective speed-up. Are there any other methods of adding basis monomials to the lookup table which take into account the degree discrepancies outlined in Section \ref{sec:deg_disc}?
\end{enumerate}

### 2022-08-02 thesis/covid19.tex
Added appendix and some tikz figures
Major grammatical and formatting changes to third draft of thesis

\begin{center}
\begin{equation*}
  \begin{split}
   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
   R_A' & = R_A + (\gamma A)\cdot \Delta \\
   R_I' & = R_I + (\gamma I)\cdot \Delta \\
   D' & = D + (\eta I)\cdot \Delta
  \end{split}
\end{equation*}
\captionof{figure}{Dynamics for the COVID19 Supermodel.}
\label{fig:covid19_dynamics}
\end{center}

where the variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}. We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate. The parameters are set based on figures shown in \cite{ansumali2020modelling}. The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$. The discretized dynamics are given in Figure \ref{fig:covid19_dynamics}.

Plots of the reachable set for this model tuned to specific values of parameters $\beta, \gamma, \eta$ were published in an ACM Sigbed Blogpost detailing applications of Formal methods for simulating disease dynamics \cite{bak2021covid}.
%
The main theme of the blogpost revolved around the difficulty of extracting accurate parameters from real-world data samples.
%
Certainly we can estimate the parameters by analyzing real-world data. However, minute changes in the parameters could yield estimates which would vastly overestimate or underestimate the true population of infected or asymptomatic patients.
%
This error is further compounded as our time horizon increases due to many of the issues pertaining to wrapping error, as discussed in the Introduction of this thesis (see Section \ref{sec:intro}).
%
Reachability analysis provides not only a method of simulating the disease dynamics in order to provide illuminating information for policy decisions but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
%
See Figure \ref{fig:covid_plots} for the two plots published in the blogpost. The Confirmed population is the sum of the number of Symptomatic ($I$) and Asymptomatic ($A$) populations of the dynamics presented in Equation \ref{eq:covid19_dynamics}. The plots were created by dividing the total period into two separate periods.
%
We decided to separate the periods as the parameters presented in the original paper \cite{ansumali2020modelling} were estimated separately according to these exact periods. Furthermore, we wished to control the conservativeness of the over-approximation of the reachable set.
%
The red line represents the real data gathered from India during the prescribed time periods with the light-blue region representing the predicted region based on the parameters given in \cite{ansumali2020modelling}.
%
Several lines representing trajectories generated according to specific parameters are also plotted in order to convey the effect of changes in the population under slight perturbations of the underlying parameters.
%
\begin{figure}
  \centering
  % First COVID plot.
  \subfloat[Confirmed population from 06/21/20-08/22/20, ]{
    \includegraphics[width= 0.9\textwidth]{figures/IndianRecoveredPlotJun21toAug22-2-2048x813.png}
  }

@@

>>>>

\begin{center}
\begin{equation*}
  \begin{cases}
   S_A' & = S_A  -(\beta S_A(A+I))\cdot \Delta \\
   S_I' & = S_I  -(\beta S_I (A + I))\cdot \Delta \\
   A' & = A + (\beta S_I(A+I) - \gamma I)\cdot \Delta \\
   I' & = I + (\beta S_I (A+I) - \gamma I)\cdot  \Delta \\
   R_A' & = R_A + (\gamma A)\cdot \Delta \\
   R_I' & = R_I + (\gamma I)\cdot \Delta \\
   D' & = D + (\eta I)\cdot \Delta
  \end{cases}
\end{equation*}
\captionof{figure}{Dynamics for the Discretized COVID19 Supermodel.}
\label{fig:covid19_dynamics}
\end{center}

The system variables denote the fraction of a population of individuals designated as \emph{Susceptible to Asymptomatic $(S_A)$}, \emph{Susceptible to Symptomatic $(S_I)$}, \emph{Asymptomatic (A)}, \emph{Symptomatic (I)}, \emph{Removed from Asymptomatic $(R_A)$}, \emph{Removed from Symptomatic $(R_I)$}, and \emph{Deceased (D)}. We choose the parameters ($\beta = 0.25, \gamma=0.02, \eta=0.02$) where $\beta$ is the probablity of infection, $\gamma$ is the removal rate, and $\eta$ is the mortality rate. The parameters are fixed based on figures shown in \cite{ansumali2020modelling}. The discretization step is chosen to be $\Delta = 0.1$ and the initial box is set to be following dimensions: $S_A  \in [0.69, 0.7], \, S_I \in [0.09, 0.1], \, A \in [0.14, 0.15], \, I \in [0.04, 0.05], \, R_A  = 0,\, R_I  = 0, \, D  = 0$. The discretized dynamics are given in Figure \ref{fig:covid19_dynamics}.

\begin{figure}[b!]
  \centering
  % First COVID plot.
  \subfloat[Confirmed population from 06/21/20-08/22/20, ]{
    \includegraphics[width= 0.9\textwidth]{figures/IndianRecoveredPlotJun21toAug22-2-2048x813.png}
  }

@@

Plots of the reachable set for this model tuned to specific values of parameters $\beta, \gamma, \eta$ were published in an ACM Sigbed Blogpost detailing applications of formal methods for simulating disease dynamics \cite{bak2021covid}.
%
The main theme revolved around the difficulty of extracting accurate parameters from real-world data samples.
%
Certainly we could attempt to estimate the parameters by analyzing real-world data. However, minute changes in the parameters could yield estimates which would vastly overestimate or underestimate the true population of infected or asymptomatic patients.
%
This error is further compounded as our time horizon increases, due to many issues pertaining to wrapping error as discussed in the introduction of this thesis (see Section \ref{sec:intro}).
%
Reachability analysis provides not only a method of simulating the disease dynamics in order to provide illuminating information for policy decisions but also to demonstrate the effect of slight changes in the parameters on the conservativeness of the outputted reachable set.
%
See Figure \ref{fig:covid_plots} for the two plots published in the blogpost. The Confirmed population is the sum of the number of Symptomatic ($I$) and Asymptomatic ($A$) populations of the dynamics presented in Equation \ref{fig:covid19_dynamics}. The plots were created by dividing the total period into two separate periods.
%
We decided to separate the periods as the parameters presented in the original paper \cite{ansumali2020modelling} were estimated separately according to these exact periods. Furthermore, we wished to control the conservativeness of the over-approximation of the reachable set.
%
The red line represents the real data gathered from India during the prescribed time periods with the light blue region representing the predicted region based on the parameters given in \cite{ansumali2020modelling}.
%
Several lines representing trajectories generated according to specific parameter values are also plotted in order to convey the effect on the confirmed population under slight perturbations of the underlying parameters.

### 2022-08-02 thesis/formats.tex
Major grammatical and formatting changes to third draft of thesis

In this chapter, we cover a method of generating template directions dynamically and automatically.
%
By dynamic, we mean that the template directions must be generated adaptively based on sampled trajectories and/or data from the state of the system.
%
By automatic, we mean that the template directions require no consideration from the user to proceed with the reachable set computation. This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation starts.
%
To briefly outline the structure of this chapter, we first expound on two techniques we utilize to dynamically generate template directions at each step.
%
The first method is based on local linear approximations where the algorithm approximates the dynamics as a linear transformation based on sample trajectories.
%
The second method is based on Principal Compoenent Analysis (PCA) where the algorithm runs PCA on the end points of the sample trajectories.
%
Finally, we cover the high-level pseudo-code of the dynamic algorithm and explain a set of parameters we feed into the algorithm in order to improve performance and the accuracy of the outputted reachable set.

@@

\section{Local Linear Approximations}
\label{sec:lin_app}
%
Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the non-linear dynamics $f$ for one step could lead to good lienar approximations of the dynamics within a small region.
%
%
To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
%
Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions over the parallelotope bundle.
%
That is the support points are the set of points $P_{supp}$ such that:
\begin{equation}
\label{eq:support}
 P_{supp} = \bigcup_{i=1}^n \{ \max_{x \in Q} \Lambda^Q_i\cdot x, \; \min_{x \in Q} \Lambda^Q_i\cdot x \}
\end{equation}
%
for all template directions of the bundle $\Lambda^Q_i$. We use the support points as a data-driven approach to find the best-fit linear function to use. These are all found by a straightforward linear program.
%
%In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
%
To find the approximate linear transformation, let $x_i$ denote the support points calculated by Equation \ref{eq:support}. We perform the following least-squares procedure: the objective would be to find an linear transformation $A$ such that it minimizes the following objective function:
\begin{equation}
  \label{eq:least_squares}
  \min_{A} \sum_{x_i} ||f(x_i) - Ax_i||^2_2
\end{equation}
%
where $||\cdot||_2$ is the standard Euclidean norm on $\reals^n$.
%
If the dynamics of a system is linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \T x \leq c_{u}$, is the set $c_{l} \leq \T \cdot A^{-1} x \leq c_{u}$.
%
Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
%

@@

\section{Principal Component Analysis}
\label{sec:pca}
The second technique for generating template directions performs Principal Component Analysis (PCA) over the images of the support points.
%
PCA is a standard technique in Statistical Machine Learning used for reduction of dimensionality by performing Singular Value Decomposition (SVD) on the covariance matrix generated by a set of data points.
%
Since the covariance matrix is symmetric (i.e $A = A^T$), the eigenvectors of this matrix will always be a orthonormal basis of the system.
%
Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
%
% Give a very light sketch of the procedure here.
To compute the images of the trajectory points, the algorithm must first compute $P_{supp}$ and propagate them one step forward by the dynamics $f$. These are the points fed into the PCA procedure.

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\section{The Dynamic Algorithm}
\label{sec:dyna_algo}
Observe that in general, our input dynamics are non-linear and therefore, the reachable set is generally non-convex.
%
On the other hand, a parallelotope bundle is always a convex set.
%
To mitigate the drawbacks of this discrepancy, we can improve accuracy of this representation by considering more template directions and more parallelotopes.
%
To this end, we add a parameter called \emph{template lifespan}, where we use the generated linear approximation and/or PCA template directions not only from the current step but also from previous steps.
%
 In our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm presented in Algorithm \ref{alg:old}.

The new approach is given in Algorithm~\ref{alg:new}.
%
During each step, the algorithm computes a collection of template directions from the two techniques outlined in Sections \ref{sec:lin_app} and \ref{sec:pca}. Each technique will be encoded as a subroutine labeled \texttt{ApproxLinearTrans} and \texttt{PCA}.
%
The \texttt{ApproxLinearTrans} function computes the best approximation of a linear transformation given a list of points before and after the one-step transformation $f$.
%
The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points.
%
Now each subroutine will return a collection of $n$ template directions, which in turn will specifiy exactly one parallelotope.
%
Hence, two parallelotopes, one generated by \texttt{ApproxLinearTrans} and the other by \texttt{PCA}, can be added to the parallelotope bundle at each step.
%

@@

Algorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
%
Line~\ref{ln:linearapprox} computes the linear approximation of the non-linear dynamics and this linear approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
%
The PCA directions of the images of support points is computed in line~\ref{ln:pca}.
%
For the time step $k$, the linear approximation and PCA templates direction matrices are given as $\Lambda_{k}^{lin}$ and $\Lambda_{k}^{pca}$, respectively.
%
To improve the accuracy of the reachable set, we compute the overapproximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the template lifespan parameter $L$.
%
Alternatively, we assign each parallelotope a parameter $L$ which dictates the number of steps we keep the parallelotope in the bundle after adding it in the current step.

>>>>

In this chapter, we cover a method of generating template directions dynamically and automatically.
%
By dynamic, we mean that the template directions must be generated adaptively based on sampled trajectories and/or data from the state of the system.
%
By automatic, we mean that the template directions require no consideration from the user to proceed with the reachable set computation. This is in contrast to the original static algorithm treated in Section \ref{sec:static} where the user must input his or her own template directions to specify the parallelotopes before the computation begans.
%
To briefly outline the structure of this chapter, we first expound on two techniques we utilize to dynamically generate template directions at each step.
%
The first method is based on local linear approximations where the algorithm approximates the dynamics as a linear transformation based on sample trajectories.
%
The second method is based on Principal Component Analysis (PCA) where the algorithm runs the PCA procedure on the image points of the sample trajectories.
%
Finally, we cover the high-level pseudo-code of the dynamic algorithm and explain a set of parameters we feed into the algorithm in order to improve performance and the accuracy of the outputted reachable set.

@@

\section{Local Linear Approximations}
\label{sec:lin_app}
%
Intuitively speaking, if time step is discretized to be sufficiently small, propagating trajectories according to the non-linear dynamics $f$ for one time step could lead to good lienar approximations of the dynamics within a small region.
%
%
To do this, we first sample a set of points in the parallelotope bundle called \emph{support points} and propagate them to the next step using the dynamics $f$.
%
Support points are a subset of the vertices of the parallelotope that either maximize or minimize the template directions over the parallelotope bundle.
%
That is the support points are the set of points $P_{supp}$ defined as:
\begin{equation}
\label{eq:support}
 P_{supp} = \bigcup_{i=1}^n \{ \max_{x \in Q} \Lambda^Q_i\cdot x, \; \min_{x \in Q} \Lambda^Q_i\cdot x \}
\end{equation}
%
for all template directions of the bundle $\Lambda^Q_i$. We use the support points as a data-driven approach to find the best-fit linear function to use. All points are found by a straightforward linear program.
%
%In this case, for small time steps a nonlinear function can be approximated fairly accurately by a linear transformation.
%
To find the approximate linear transformation, let $x_i$ denote the support points calculated by Equation \ref{eq:support}. We perform the following least-squares procedure: the objective would be to find an linear transformation $A$ such that it minimizes the following objective function:
\begin{equation}
  \label{eq:least_squares}
  \min_{A} \sum_{x_i} ||f(x_i) - Ax_i||^2_2
\end{equation}
%
where $||\cdot||_2$ is the standard Euclidean norm on $\reals^n$.
%
If the dynamics of a system are linear, i.e., $x^{+} = Ax$, the image of the parallelotope $c_{l} \leq \Lambda x \leq c_{u}$, is the set $c_{l} \leq \Lambda \cdot A^{-1} x \leq c_{u}$.
%
Therefore, given the template directions of the initial set as $\T_0$, we compute the local linear approximation of the non-linear dynamics and change the template directions by multiplying them with the inverse of the approximate linear dynamics.
%
%This will reflect the approximate template directions for one step forward. To continue to
%

@@

\section{Principal Component Analysis}
\label{sec:pca}
The second technique for generating template directions performs Principal Component Analysis (PCA) over the images of the support points.
%
PCA is a standard technique in Statistical Machine Learning used to reduction of dimensionality by performing Singular Value Decomposition (SVD) on the covariance matrix generated by a set of data points.
%
Since the covariance matrix is symmetric by definition (i.e $A = A^T$), the eigenvectors of this matrix will always be a orthonormal basis of the system.
%
Using PCA is a reasonable choice as it produces orthonormal directions that can construct a rotated box for bounding the points.
%
% Give a very light sketch of the procedure here.
To compute the images of the trajectory points, the algorithm must first compute $P_{supp}$ and propagate them one step forward by the dynamics $f$. These image points fed into the PCA procedure.

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\section{The Dynamic Algorithm}
\label{sec:dyna_algo}
Observe that in general, our input dynamics are non-linear and therefore, the reachable set is generally non-convex.
%
On the other hand, a parallelotope bundle is always a convex set.
%
To mitigate the drawbacks of this discrepancy, we can improve the accuracy of this representation by considering more template directions and more parallelotopes.
%
To this end, we add a parameter called \emph{template lifespan}, where we use the generated linear approximation and/or PCA template directions not only from the current step but also from previous steps.
%
During our benchmarks, we tune each of the options (PCA / linear approximation as well as lifespan option) to demonstrate that specific parameters generate more accurate reachable sets than those generated by the static algorithm presented in Algorithm \ref{alg:old}.

The new approach is given in Algorithm~\ref{alg:new}.
%
During each step, the algorithm computes a collection of template directions from the two techniques outlined in the previous two sections. The subroutines will be encoded as a subroutine labeled \texttt{ApproxLinearTrans} and \texttt{PCA} respectively.
%
The \texttt{ApproxLinearTrans} function computes the best linear approximation of the dynamics as specified in Section \ref{sec:lin_app}.
%
The \texttt{PCA} function returns a set of orthogonal directions using principal component analysis of a set of points as specified in Section \ref{sec:pca}.
%
Now each subroutine will return a collection of $n$ template directions, which in turn will specifiy exactly one parallelotope.
%
Hence, two parallelotopes, one generated by \texttt{ApproxLinearTrans} and the other by \texttt{PCA}, can be added to the parallelotope bundle at each step.
%

@@

Algorithm~\ref{alg:new} computes the dynamic templates for each time step $k$.
%
Line~\ref{ln:linearapprox} computes the linear approximation of the non-linear dynamics and this approximation is used to compute the new template directions according to this linear transformation in Line~\ref{ln:linapp}.
%
The PCA directions of the images of support points is computed in Line~\ref{ln:pca}.
%
For the time step $k$, the linear approximation and PCA templates direction matrices are given as $\Lambda_{k}^{lin}$ and $\Lambda_{k}^{pca}$, respectively.
%
To improve the accuracy of the reachable set, we compute the over-approximation of the reachable set with respect to not just the template directions at the current step, but with respect to other template directions for time steps that are within the template lifespan parameter $L$.
%
Alternatively, we assign each parallelotope a parameter $L$ which dictates the number of steps we keep the parallelotope in the bundle after adding it in the current step.

### 2022-08-02 thesis/frontmatter/abstract.tex
Major grammatical and formatting changes to third draft of thesis

\end{center}
\par Reachable set computation is an important technique for the verification of safety properties of dynamical systems.
%
% one method proposed for solving this problem, based on parallelotope bundle reachability for
%
In this thesis, we investigate reachable set computation for discrete nonlinear systems based on parallelotope bundles.
%
The crux of the reachability algorithm relies on computing an upper and lower bound on the supremum and infimum respectively of a nonlinear function over a rectangular domain.
%
We cover two ways of computing these bounds: one method utilizing Bernstein polynomials and the other relying on a non-linear optimization tool developed by NASA, Kodiak.
%
We aim to improve the traditional parallelotope-based reachability method by removing the manual step of parallelotope template selection in order to make the procedure fully automatic.
%
Furthermore, we show that adding templates dynamically during computations can improve accuracy.
%
% One of the simplest algorithms for computing reachable sets for discrete nonlinear systems uses parallelotope bundles and Bernstein polynomials.
%
%Our main hypothesis in this paper is that generating templates in a dynamic manner would improve the accuracy of the reachable set.
%
To this end, we investigate two techniques for generating the template directions.
%
The first technique approximates the dynamics as a linear transformation and generates templates using this linear transformation.
%
The second technique uses Principal Component Analysis (PCA) of sample trajectories for generating templates.
%
We have implemented our approach in a Python-based tool called Kaa.
%
%Kaa was developed as a more compact, simple implementation of the pre-existing parallelotope-based reachability tool, Sapo.
%
% have implemented two main enhancements to it.
%
The tool is modular and use two types of global optimization solvers, the first using Bernstein polynomials and the second using
the aforementioned Kodiak library.
%
We demonstrate the improved accuracy of our approach on several standard nonlinear benchmark systems, including a high-dimensional COVID19 model proposed by the Indian Supermodel Committee.
%
Finally, we explore a particular potential application of the Bernstein expansion technique to real-time reachability.
%
We present evidence of several hurdles and barriers against effectively utilizing our Bernstein coefficient pruning method.
\clearpage

>>>>

\end{center}
\par Reachable set computation is an important technique for the verification of safety properties of dynamical systems.
%
% one method proposed for solving this problem, based on parallelotope bundle reachability for
%
In this thesis, we investigate reachable set computation for discrete non-linear systems based on parallelotope bundles.
%
The crux of the reachability algorithm relies on computing an upper and lower bound on the supremum and infimum respectively of a non-linear function over a rectangular domain.
%
Bernstein Expansion of a polynomial function has been explored as a traditional method for computing these bounds effciently.
%
In light of this, we aim to improve the traditional parallelotope-based reachability method by removing the manual step of parallelotope template selection in order to make the procedure fully automatic.
%
Furthermore, we show that adding templates dynamically during computations can improve accuracy.
%
To this end, we investigate two techniques for generating template directions.
%
The first technique approximates the dynamics as a linear transformation and generates templates using this transformation.
%
The second technique uses Principal Component Analysis (PCA) of sample trajectories for generating templates.
%
We have implemented our approach in a Python-based tool called Kaa, which uses two types of global optimization solvers, the first using Bernstein polynomials and the second using
the Kodiak library.
%
We demonstrate the improved accuracy of our approach on several standard nonlinear benchmark systems, including a high-dimensional COVID19 model.
%
Finally, we explore a potential application of the Bernstein expansion technique to real-time reachability.
%
We present evidence of several hurdles and barriers against effectively utilizing our Bernstein coefficient pruning method.
\clearpage

### 2022-08-02 thesis/frontmatter/acknowledgement.tex
Major grammatical and formatting changes to third draft of thesis

I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community for pursuing this research. In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
%
%His choice to give me more autonomy to take classes at Duke University really helped me understand my academic interests better.
%
Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proofreading drafts, and socializing over group lunches. I would also like to thank Stanley Bak for helping me improve the implementation of Kaa and providing valuable feedback during this research. Juan Garcia made a wonderful collegue and an equally wonderful friend. The nights we raided Peabody and Phillips hall was certainly a memorable time. Finally, I would like to thank my parents and my family for their ardent, un-ending support.

>>>>

I would like to thank the UNC Computer Science community for supporting my work and providing me a warm community to pursue this research. In particular, I would like to thank my advisor, Parasara Sridhar Duggirala, for the opporunity to perform research and participate in his group.
%
%His choice to give me more autonomy to take classes at Duke University really helped me understand my academic interests better.
%
Bineet Ghosh, Manish Goyal, Abel Karimi, and Meghan Stuart were all great colleagues and I enjoyed my time exchanging ideas, proof-reading drafts, and socializing over group lunches. I would also like to thank Stanley Bak for helping me improve the implementation of Kaa and providing valuable feedback during this research. Juan Garcia is a wonderful collegue and an equally wonderful friend. The nights we raided Peabody and Phillips hall was certainly a memorable time. Finally, I would like to thank my parents and my family for their ardent, unending support.
